---
title: Assessment of the relationship between the type of the front of the car and   fuel consumption
author: "ps7391"
date: '7 January 2019'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Executive Summary

Consider a dataset of a collection of cars, and interested in exploring the relationship between a set of variables and miles per gallon. In particularly the following two questions:

1. Is an automatic or manual transmission better for MPG
2. Quantify the MPG difference berween automatic and manual transmissions

Perform the following sequence of actions:

1. Process the data, for use of this project
2. Explore the data, especially focussing on the two paramaters we are interested in transmission and MPG
3. Model selection, where we try different models to help answer our questions
4. Model examination, to see wether our best model holds up to our standards
5. Conclusion where we answer the questions based on the data

### Processing

In the first place change 'am' to factor (0 = automatic, 1 = manual) and make cylinders a factor as well (since it is not continious)


```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
library(GGally)
library(dplyr)
library(ggfortify)

data("mtcars")

mtcars_factors <- mtcars
mtcars_factors$am <- as.factor(mtcars_factors$am)
levels(mtcars_factors$am) <- c("automatic", "manual")

mtcars_factors$cyl <- as.factor(mtcars_factors$cyl)
mtcars_factors$gear <- as.factor(mtcars_factors$gear)
mtcars_factors$vs <- as.factor(mtcars_factors$vs)
levels(mtcars_factors$vs) <- c("V", "S")
```

### Exploratory data analyses

Look at the dimensions and head of the dataset:

```{r}
# dimensions
dim(mtcars_factors)

# head
head(mtcars_factors)
```

Find the relationship between the two parameters of interest.

```{r}
g <- ggplot(mtcars_factors, aes(x = am, y = mpg, fill = am))
g + geom_boxplot() +
        scale_fill_manual(name = "am", values = c("yellow", "green")) +
        ggtitle("Relationship between MPG and automatic or manual geer") + 
        theme(plot.title = element_text(face = "bold", size = 12))
```

